### 2023-11-27 R_code/paper_analysis.Rmd
fix

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
library(scales)
library(tidyverse)
library(caret)
library(ggpubr)
library(pROC)
library(predtools)
library(MASS)
library(caret)
```

```{r}
prefix = "/Users/noa/Workspace/bootstrap_results/remote_results/"

fasttree_all_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/all_models_performance.tsv", sep = ""))
fasttree_final_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/final_model_performance.tsv", sep = ""))
fasttree_group_performance = read_tsv(paste(prefix,"ML_results/fasttree/groups_performance.tsv", sep = ""))
fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/model_standard_vi.tsv", sep = ""))
fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/final_model/test.tsv", sep = ""))
fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
fasttree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
fasttree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
fasttree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))
#####
iqtree_all_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/all_models_performance.tsv", sep = ""))
iqtree_final_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/final_model_performance.tsv", sep = ""))
iqtree_group_performance = read_tsv(paste(prefix,"ML_results/iqtree/groups_performance.tsv", sep = ""))
iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/model_standard_vi.tsv", sep = ""))
iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/final_model/test.tsv", sep = ""))
iqtree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_downgrade.tsv", sep = ""))
iqtree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_upgrade.tsv", sep = ""))
iqtree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_standard.tsv", sep = ""))
#####
raxml_all_models_performance = read_tsv(paste(prefix,"ML_results/raxml/all_models_performance.tsv", sep = ""))
raxml_final_models_performance = read_tsv(paste(prefix,"ML_results/raxml/final_model_performance.tsv", sep = ""))
raxml_group_performance = read_tsv(paste(prefix,"ML_results/raxml/groups_performance.tsv", sep = ""))
raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/model_standard_vi.tsv", sep = ""))
raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml/final_model/test.tsv", sep = ""))
raxml_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
raxml_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
raxml_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))

```

General performance
IQTREE
```{r}

out_path <-paste(prefix,"ML_results/","iqtree","/performance_text",".tsv", sep = "")

paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,brier_loss,logloss, AUC,average_precision,mcc_0.5, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ ultrafast bootstrap support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description, brier_loss,logloss,AUC,average_precision, mcc_0.5,FPR,FNR, )  %>% arrange(dataset)

paper_performance_iqtree
write_tsv(paper_performance_iqtree,out_path)

paper_performance_iqtree %>% arrange(description)

```
FASTTREE
```{r}

out_path <-paste(prefix,"ML_results/","fasttree","/performance_text",".tsv", sep = "")

paper_performance_fasttree<-fasttree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw SH support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated SH support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ SH support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)

paper_performance_fasttree
write_tsv(paper_performance_fasttree,out_path)
paper_performance_fasttree %>% arrange(description)

```
RAXML
```{r}

out_path <-paste(prefix,"ML_results/","raxml","/performance_text",".tsv", sep = "")

paper_performance_raxml<-raxml_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw bootstrap support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated bootstrap support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ bootstrap support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)

paper_performance_raxml
write_tsv(paper_performance_raxml,out_path)
paper_performance_raxml %>% arrange(description)

```

Results on test data
```{r}
total_model_performance = rbind(paper_performance_iqtree %>% mutate(program='IQTREE'), paper_performance_fasttree %>% mutate(program='FastTree'), paper_performance_raxml %>% mutate(program='RAxML-NG')) %>% filter (dataset %in% c("test"), description %in% c("all features","Raw SH support","Raw aLRT support","Raw aBayes support","Raw bootstrap support","Raw ultrafast bootstrap support")) %>% mutate(Support = description) %>% rename(MCC = mcc_0.5)

mcc_plot<-total_model_performance %>% ggplot(aes(y=MCC, x = program, fill = Support))+geom_col(position = "dodge")
logloss_plot<-total_model_performance %>% ggplot(aes(y=logloss, x = program, fill = Support))+geom_col(position = "dodge")
#total_model_performance %>% ggplot(aes)

ggarrange(mcc_plot,logloss_plot, labels = c("A","B"),align = "h", nrow = 2, ncol = 1, common.legend = TRUE )
```

Variable importance 
IQTREE

```{r}
out_path <-paste(prefix,"ML_results/","iqtree","/vi_text",".tsv", sep = "")

vi_data<- iqtree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',

   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",

`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`,feature_name)
vi_data
write_tsv(vi_data,out_path)

```

FASTTREE

```{r}
out_path <-paste(prefix,"ML_results/","fasttree","/vi_text",".tsv", sep = "")

vi_data<- fasttree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',

   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",

`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
vi_data 
write_tsv(vi_data,out_path)
```

RAXML

```{r}
out_path <-paste(prefix,"ML_results/","raxml","/vi_text",".tsv", sep = "")

vi_data<- raxml_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',

   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",

`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
vi_data %>% arrange(feature_name)
write_tsv(vi_data,out_path)
```

Confusion matrix

```{r}
ggplotConfusionMatrix <- function(m){
  mytitle <- ""#paste("Accuracy", percent_format()(m$overall[1]))
                  # "Kappa", percent_format()(m$overall[2]))
  p <-
    ggplot(data = as.data.frame(m$table) ,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = scales::comma(Freq))) +
    theme(legend.position = "none") +
    ggtitle(mytitle)+theme(text = element_text(size = 13))
  return(p)
}
conf_data<-iqtree_enriched_test %>% dplyr::select (prob_predictions, true_label) %>% mutate (pred_status = prob_predictions>0.5)

obs<- factor(conf_data %>% pull (true_label))
pred<- factor(conf_data %>% pull (pred_status))

confusion_plot<-ggplotConfusionMatrix(confusionMatrix(data = pred, reference= obs))+theme(legend.position = "none")

error_vs_size<- all_models_performance %>% filter (metric_type=='all_data', dataset =='test', analysis_type=='full_standard')
error_vs_size_plt = error_vs_size %>% mutate(n_MSAs = sample_frac*2007) %>% ggplot(aes(x=n_MSAs, y = mcc))+geom_line()+ expand_limits(y = 0)+xlab("Number of MSAs")+ theme(text = element_text(size = 13))+ylab("MCC")+geom_point()

rocobj <- roc(enriched_test$true_label, enriched_test$prob_predictions)
auc <- round(auc(enriched_test$true_label, enriched_test$prob_predictions),3)

global_max_auc_plt<-ggroc(rocobj,colour = 'steelblue', size = 2)+
  ggtitle(paste0('AUC = ', auc))+theme(text = element_text(size = 13))+xlab("Specificity")

ggarrange(global_max_auc_plt,confusion_plot,error_vs_size_plt, labels = c("A","B","C"),align = "h", nrow = 2, ncol = 2 )

```

Group performance

```{r}
test_group_performance<- group_performance %>% filter (dataset=='test')

Pypythia_error<-test_group_performance %>% filter (grouping_col_name =='msa_difficulty_group') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(0.00328, 0.303]','(0.303, 0.538]','(0.538, 0.571]','	(0.571, 0.865]'))) %>%
ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('MSA\ndifficulty')+ xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))

n_seq_error<-test_group_performance %>% filter (grouping_col_name =='n_seq_group')   %>%  mutate(grouping_col = factor(grouping_col, levels = c('(4.999, 69.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]'))) %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof sequences')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))

n_loci_error<-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(61.999, 727.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]')))  %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof positions')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))

Pypythia_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions,  feature_msa_pypythia_msa_difficulty) %>% group_by(msa_path,feature_msa_pypythia_msa_difficulty) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty, y = mean_prob))+geom_point(size = 0.7)+xlab("MSA\ndifficulty")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))

seq_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions, feature_msa_n_seq, feature_msa_pypythia_msa_difficulty) %>% group_by(msa_path,feature_msa_n_seq) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_n_seq, y = mean_prob))+geom_point(size = 0.7)+xlab("Number of\nsequences")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))

n_loci_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions, feature_msa_n_loci) %>% group_by(msa_path,feature_msa_n_loci) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_n_loci, y = mean_prob))+geom_point(size = 0.7)+xlab("Number of\n positions")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))

ggarrange(seq_scatter,Pypythia_scatter,n_loci_scatter, n_seq_error,Pypythia_error, n_loci_error,labels = c("A","C","E","B","D","F"),align = "h", nrow = 3, ncol = 3, legend = "bottom",vjust= 1)
#error_vs_group_5_5_including_plt

cor.test( enriched_test$feature_msa_pypythia_msa_difficulty,  enriched_test$prob_predictions)
cor.test( enriched_test$feature_msa_n_seq, enriched_test$prob_predictions)
cor.test( enriched_test$feature_msa_n_loci, enriched_test$prob_predictions)

```

Running time

```{r}
running_time_data<- enriched_test%>% dplyr::select (boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = boot_run_time/extraction_running_time, diff2 = boot_run_time/(extraction_running_time+nni_ll_eval_running_time
)) 
summary(running_time_data %>% pull (diff1))
summary(running_time_data %>% pull (diff2))
summary(running_time_data %>% pull (extraction_running_time))
summary(running_time_data %>% pull (nni_ll_eval_running_time))
#enriched_test %>% dplyr::select (boot_run_time, extraction_runni
```

```{r}

running_time_data<- enriched_test%>% dplyr::select (program_boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
)) 
summary(running_time_data %>% pull (diff1))
summary(running_time_data %>% pull (diff2))
summary(running_time_data %>% pull (extraction_running_time))
summary(running_time_data %>% pull (nni_ll_eval_running_time))
summary(running_time_data %>% pull (program_boot_run_time))
#enriched_test %>% dplyr::select (boot_run_time, extraction_running_time, nni_ll_eval_running_time)
```

```{r}
summary(validation_data$feature_n_unique_seq)
raxml_validation_data$
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.