### 2023-11-30 R_code/paper_analysis.Rmd
new changes for ML pipeline

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
library(scales)
library(tidyverse)
library(ggpubr)
library(pROC)
library(predtools)
library(MASS)
```

```{r}
prefix = "/Users/noa/Workspace/bootstrap_results/remote_results/"

fasttree_all_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/all_models_performance.tsv", sep = ""))
fasttree_final_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/final_model_performance.tsv", sep = ""))
fasttree_group_performance = read_tsv(paste(prefix,"ML_results/fasttree/groups_performance.tsv", sep = ""))
fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/model_standard_vi.tsv", sep = ""))
fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/final_model/test.tsv", sep = ""))
fasttree_enriched_train = read_tsv(paste(prefix,"ML_results/fasttree/final_model/train.tsv", sep = ""))
fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
fasttree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
fasttree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
fasttree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))
#####
iqtree_all_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/all_models_performance.tsv", sep = ""))
iqtree_final_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/final_model_performance.tsv", sep = ""))
iqtree_group_performance = read_tsv(paste(prefix,"ML_results/iqtree/groups_performance.tsv", sep = ""))
iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/model_standard_vi.tsv", sep = ""))
iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/final_model/test.tsv", sep = ""))
iqtree_enriched_train = read_tsv(paste(prefix,"ML_results/iqtree/final_model/train.tsv", sep = ""))
iqtree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_downgrade.tsv", sep = ""))
iqtree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_upgrade.tsv", sep = ""))
iqtree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_standard.tsv", sep = ""))
#####
raxml_all_models_performance = read_tsv(paste(prefix,"ML_results/raxml/all_models_performance.tsv", sep = ""))
raxml_final_models_performance = read_tsv(paste(prefix,"ML_results/raxml/final_model_performance.tsv", sep = ""))
raxml_group_performance = read_tsv(paste(prefix,"ML_results/raxml/groups_performance.tsv", sep = ""))
raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/model_standard_vi.tsv", sep = ""))
raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml/final_model/test.tsv", sep = ""))
raxml_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
raxml_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
raxml_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))

```

```{r}
iqtree_enriched_train %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
iqtree_enriched_test %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
iqtree_enriched_val_standard %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
iqtree_enriched_val_upgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
iqtree_enriched_val_downgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()

iqtree_enriched_test %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
iqtree_enriched_val_standard %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
iqtree_enriched_val_upgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
iqtree_enriched_val_downgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
```

Results on test data
```{r}
total_model_performance = rbind(paper_performance_iqtree %>% mutate(program='IQTREE'), paper_performance_fasttree %>% mutate(program='FastTree'), paper_performance_raxml %>% mutate(program='RAxML-NG')) %>% filter (dataset %in% c("test"), description %in% c("all features","Raw SH support","Raw aLRT support","Raw aBayes support","Raw bootstrap support","Raw ultrafast bootstrap support")) %>% mutate(Support = description) %>% rename(MCC = mcc_0.5)

mcc_plot<-total_model_performance %>% ggplot(aes(y=MCC, x = program, fill = Support))+geom_col(position = "dodge")
logloss_plot<-total_model_performance %>% ggplot(aes(y=logloss, x = program, fill = Support))+geom_col(position = "dodge")
#total_model_performance %>% ggplot(aes)

ggarrange(mcc_plot,logloss_plot, labels = c("A","B"),align = "h", nrow = 2, ncol = 1, common.legend = TRUE )
```

Variable importance 
IQTREE

```{r}
out_path <-paste(prefix,"ML_results/","iqtree","/vi_text",".tsv", sep = "")

vi_data<- iqtree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',

   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",

`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`,feature_name)
vi_data
write_tsv(vi_data,out_path)

```

FASTTREE

```{r}
out_path <-paste(prefix,"ML_results/","fasttree","/vi_text",".tsv", sep = "")

vi_data<- fasttree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',

   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",

`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
vi_data 
write_tsv(vi_data,out_path)
```

RAXML

```{r}
out_path <-paste(prefix,"ML_results/","raxml","/vi_text",".tsv", sep = "")

vi_data<- raxml_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",

  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',

   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",

`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",

`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',

`...1` =='feature_partition_branch'~'Branch length at the parititon',
`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
`...1` =='feature_mean_bl'~'Average of tree branch length',

`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',

`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',

)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
vi_data 
write_tsv(vi_data,out_path)
```

Performance per sample fraction

```{r}
iqtree_data<- iqtree_all_models_performance %>% filter (dataset=='test', name =='model_standard', analysis_type=='full_standard')  %>% dplyr::select (sample_frac, AUC,logloss) %>% mutate (Program = 'IQTREE')

raxml_data<- raxml_all_models_performance %>% filter (dataset=='test', name =='model_standard', analysis_type=='full_standard')  %>% dplyr::select (sample_frac, AUC,logloss) %>% mutate (Program = 'RAxML-NG')

fasttree_data<- fasttree_all_models_performance %>% filter (dataset=='test', name =='model_standard', analysis_type=='full_standard')  %>% dplyr::select (sample_frac, AUC,logloss) %>% mutate (Program = 'FastTree')

sample_frac_vs_result_plot<-rbind(iqtree_data,raxml_data,fasttree_data) %>% ggplot(aes(x=sample_frac*4652, y = logloss, color = Program))+geom_line()+expand_limits(y = c(0,0))+xlab("Number of training MSAs")+theme_classic()
sample_frac_vs_result_plot
```

Performance per group

```{r}
library(yardstick)
per_group_data_iqtree<-iqtree_enriched_test%>% dplyr::select (predictions_full_standard, true_binary_support, model_short, feature_msa_pypythia_msa_difficulty,tree_search_model,feature_msa_n_loci, tree_search_model,feature_msa_n_seq) %>% mutate(Pypythia_group = ntile(feature_msa_pypythia_msa_difficulty,n = 20),
n_seq_group  = ntile(feature_msa_n_seq, 20), n_loci_group = ntile(feature_msa_n_loci,n = 20), true = as.factor(true_binary_support), pred = 1-predictions_full_standard) %>% mutate(Program='IQTREE') %>% group_by(n_seq_group) %>% mutate(median_seq_per_group = median(feature_msa_n_seq)) %>% ungroup() %>% group_by(n_loci_group) %>% mutate(median_loci_per_group = median(feature_msa_n_loci)) %>% ungroup() %>% group_by(Pypythia_group) %>% mutate(median_pythia_per_group = median(feature_msa_pypythia_msa_difficulty)) %>% ungroup()

per_group_data_fasttree<-fasttree_enriched_test%>% dplyr::select (predictions_full_standard, true_binary_support, model_short, feature_msa_pypythia_msa_difficulty,tree_search_model,feature_msa_n_loci, tree_search_model,feature_msa_n_seq) %>% mutate(Pypythia_group = ntile(feature_msa_pypythia_msa_difficulty,n = 20),
n_seq_group  = ntile(feature_msa_n_seq, 20), n_loci_group = ntile(feature_msa_n_loci,n = 20), true = as.factor(true_binary_support), pred = 1-predictions_full_standard) %>% mutate(Program='FastTree') %>% group_by(n_seq_group) %>% mutate(median_seq_per_group = median(feature_msa_n_seq)) %>% ungroup() %>% group_by(n_loci_group) %>% mutate(median_loci_per_group = median(feature_msa_n_loci)) %>% ungroup() %>% group_by(Pypythia_group) %>% mutate(median_pythia_per_group = median(feature_msa_pypythia_msa_difficulty)) %>% ungroup()

per_group_data_raxml<-raxml_enriched_test%>% dplyr::select (predictions_full_standard, true_binary_support, model_short, feature_msa_pypythia_msa_difficulty,tree_search_model,feature_msa_n_loci, tree_search_model,feature_msa_n_seq) %>% mutate(Pypythia_group = ntile(feature_msa_pypythia_msa_difficulty,n = 20),
n_seq_group  = ntile(feature_msa_n_seq, 20), n_loci_group = ntile(feature_msa_n_loci,n = 20), true = as.factor(true_binary_support), pred = 1-predictions_full_standard) %>% mutate(Program='RAxML-NG') %>% group_by(n_seq_group) %>% mutate(median_seq_per_group = median(feature_msa_n_seq)) %>% ungroup() %>% group_by(n_loci_group) %>% mutate(median_loci_per_group = median(feature_msa_n_loci)) %>% ungroup() %>% group_by(Pypythia_group) %>% mutate(median_pythia_per_group = median(feature_msa_pypythia_msa_difficulty)) %>% ungroup()

total_data <- rbind(per_group_data_iqtree, per_group_data_fasttree, per_group_data_raxml)

pythia<- total_data  %>% group_by(median_pythia_per_group, Program) %>% roc_auc( true, pred) %>% mutate(AUC=.estimate) %>% ggplot(aes(x=median_pythia_per_group, y = AUC, color = Program))+expand_limits(y = c(0.5,1))+geom_line()+xlab("MSA difficulty")+ theme_classic()
loci<-total_data %>% group_by(median_loci_per_group,Program) %>% roc_auc( true, pred) %>% mutate(AUC=.estimate) %>% ggplot(aes(x=median_loci_per_group, y = AUC, color = Program))+expand_limits(y = c(0.5,1))+geom_line()+xlab("Number of positions")+ theme_classic()
n_seq<-total_data  %>% group_by(median_seq_per_group, Program) %>% roc_auc( true, pred) %>% mutate(AUC=.estimate) %>% ggplot(aes(x=median_seq_per_group, y = AUC, color = Program))+expand_limits(y = c(0.5,1))+geom_line()+xlab("Number of sequences")+ theme_classic()

```

```{r}
ggarrange(sample_frac_vs_result_plot,pythia,loci,n_seq, labels = c("A","B","C","D"),align = "h", nrow = 2, ncol = 2, common.legend = TRUE )
```

```{r}
cor.test(iqtree_enriched_test$predictions_full_standard, iqtree_enriched_test$feature_msa_pypythia_msa_difficulty)
cor.test(raxml_enriched_test$predictions_full_standard, raxml_enriched_test$feature_msa_pypythia_msa_difficulty)
cor.test(fasttree_enriched_test$predictions_full_standard, fasttree_enriched_test$feature_msa_pypythia_msa_difficulty)

cor.test(iqtree_enriched_test$predictions_full_standard, iqtree_enriched_test$feature_msa_n_loci)
cor.test(raxml_enriched_test$predictions_full_standard, raxml_enriched_test$feature_msa_n_loci)
cor.test(fasttree_enriched_test$predictions_full_standard, fasttree_enriched_test$feature_msa_n_loci)

cor.test(iqtree_enriched_test$predictions_full_standard, iqtree_enriched_test$feature_msa_n_seq)
cor.test(raxml_enriched_test$predictions_full_standard, raxml_enriched_test$feature_msa_n_seq)
cor.test(fasttree_enriched_test$predictions_full_standard, fasttree_enriched_test$feature_msa_n_seq)

```

```{r}
iqtree_val
```

```{r}

running_time_data<- raxml_enriched_test%>% dplyr::select (raxml_enriched_test$time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
)) 
summary(running_time_data %>% pull (diff1))
summary(running_time_data %>% pull (diff2))
summary(running_time_data %>% pull (extraction_running_time))
summary(running_time_data %>% pull (nni_ll_eval_running_time))
summary(running_time_data %>% pull (program_boot_run_time))
#enriched_test %>% dplyr::select (boot_run_time, extraction_running_time, nni_ll_eval_running_time)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.